---
title: "Homework 5"
author: Nikhita Damaraju
output: github_document
---

This is my solution to Homework 5.

Library imports 

```{r settings, message = FALSE} 
library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)


theme_set(theme_bw() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
homicide_df = 
  read_csv('./problem1_data/homicide-data.csv')
```


## Problem 2

Making combined tidy dataset using all the files 

```{r message = FALSE}
logitudinal_df = 
  tibble(
    path = list.files("./problem2_data/"),
  ) %>%
  mutate(
    path = str_c("./problem2_data/", path),
    data = map(.x = path, ~ read_csv(.x)),
    arm = case_when(
      str_detect(path, 'con') ~ 'control',
      str_detect(path, 'exp') ~ 'experimental'
    ),
    subject_id = str_sub(path, start = -6L, end = -5L)
  ) %>%
  select(-path) %>%
  unnest(data) 
```

Spaghetti plot showing observations over time per group

```{r}
logitudinal_df %>%
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    names_prefix = "week_",
    values_to = "observation"
  ) %>%
  ggplot(aes(x = week, y = observation, group = subject_id, color = subject_id)) +
  geom_line() +
  facet_grid(. ~ arm) + 
  labs(
    title = "Observation for each subject over time",
    x = "Week number",
    y = "Observation"
  )
  
```

The data from the longitudinal study contains 10 participants with unique subject IDs that have observations from a control arm and experimental arm over a span of 8 weeks. It can be observed that the observation values for the experimental arm are higher that the control arm. Thus, there is a visual difference between the two groups. 

## Problem 3

Function to calculate all required variables for each dataset

```{r}
sim_hypothesis = function(n, mu, sigma = 5) {
  
  sim_data = tibble(
    samp = rnorm(n, mean = mu, sd = sigma),
  )
  
  sim_data %>%
    summarize(
      mu_hat = mean(samp),
      mu = mu,
      t.test(samp, mu = 0, alternative = 'two.sided', paired = FALSE, conf.level = 0.95) %>% broom::tidy() %>% select(p.value)
   ) 
  
}
  
```

Running function for different values of mu for 5000 iterations

```{r}
output = vector("list", 7)

for (i in 1:7) {
  output[[i]] = rerun(5000, sim_hypothesis(30, (i - 1))) %>%
    bind_rows()
}

sim_results = bind_rows(output) 
```

Proportion of times null was rejected

```{r}
sim_results =
  sim_results %>%
  mutate(
    test_result = case_when(
      p.value < .05 ~ "reject",
      p.value >= .05 ~ "fail to reject"
    ))

sim_results %>%
  filter(test_result == 'reject') %>%
  group_by(mu) %>%
  summarize(
    proportion = n()/5000
  ) %>%
  ggplot(aes(x = mu, y = proportion)) +
  geom_point() +
  geom_line() +
  labs(
    title = "proportion of null hypothesis rejects vs true mean",
    x = "true mean (mu)",
    y = "proportion of times null hypothesis was rejected"
  )
```


